---
title: "Project_Proposal"
author: "Tim Eckart"
date: "2026-07-12"
output:
  html_document:
    theme: readable
    toc: true
    highlight: zenburn
  pdf_document:
    toc: true
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Proposal

*1.)* Load your dataset into R and take a snapshot of the data. There is no requirement on how much to show, and there is no need to show the entire dataset. Even 10 rows should be enough.

```{r Proposal_1a}
library(data.table)
library(readr)
azure_vm_raw = read_csv("trace_data_vmtable_vmtable.csv", col_names = FALSE, n_max = 10) 
azure_vm_raw
```
- **Answer:** Here is the data set listed from Microsoft Azure Public Dataset V2 on Github: https://github.com/Azure/AzurePublicDataset/releases/tag/dataset-v2 . The data vmtable file has no header rows. The names are from the repository's schema file. Here is also a breakdown of the headers: https://github.com/Azure/AzurePublicDataset/blob/master/AzurePublicDatasetV2.md 

```{r Proposal_1b}
azure_schema = read_csv("schema.csv")
knitr::kable(azure_schema,
             caption = "Schema for the Azure 2019 VM trace (from schema.csv)")
```
- Here is the Schema.csv listing the column headers name for the table.

```{r Proposal_1c}
library(readr)

vm_cols = c("vm_id", "subscription_id", "deployment_id", "timestamp_vm_created", "timestamp_vm_deleted", "max_cpu", 
            "avg_cpu", "p95_max_cpu", "vm_category", "vm_virtual_core_count_bucket", "vm_memory_gb_bucket")

azure_vm_table = read_csv("trace_data_vmtable_vmtable.csv", col_names = vm_cols, 
                          col_types = cols(vm_virtual_core_count_bucket = col_character(), vm_memory_gb_bucket = col_character()))

head(azure_vm_table[, -c(4:11)],10)

head(azure_vm_table[, -c(1:3)],10)
head(azure_vm_table[, -c(1:8)],10)

```
- Here is the columns renamed to their respected headers per Schema. 

**2.)**	Write a paragraph saying who is working on the project (if team), or your name otherwise. Explain briefly what the dataset is about. Provide a link so I can access the dataset if necessary. State how many observations there are.\n

Finally, write one or two sentences about what the goal of the analysis is. Clearly state the response variable and explanatory variables. This should be a simple thing such as: The goal of this analysis is to predict the mpg for certain cars based on x,y,z,a,b,c,…\n\n

- **Answer:** This project will be conducted individually by Timothy Eckart, using the **vmtable trace** from the **Azure Public Dataset V2**, published by Microsoft in the AzurePublicDataset repository on GitHub (https://github.com/Azure/AzurePublicDataset/releases/tag/dataset-v2). The dataset is a sanitized, representative subset of Microsoft Azure's first-party virtual machine workload in one geographical region, collected over 30 consecutive days in 2019 and released under a CC-BY license. The data contains 2,695,548 observations on virtual machines from 6,687 Azure customer subscriptions, including average, maximum, and 95th-percentile CPU utilization, VM category, virtual core count, memory size, and creation/deletion time stamps. Each observation represents one virtual machine's full lifetime within the trace. The goal of this analysis is to predict a virtual machine's average CPU utilization (`avg_cpu`) based on its category, virtual core count, memory size, and lifetime derived from its creation and deletion timestamp. \n

- Per the dataset's CC-BY attribution requirement this project cites the accompanying SOSP’17 paper (Cortez 2017; see References).

## Reference

Cortez, E., Bonde, A., Muzio, A., Russinovich, M., Fontoura, M., & Bianchini, R. (2017). Resource Central: Understanding and predicting
workloads for improved resource management in large cloud platforms. In *Proceedings of the 26th Symposium on Operating Systems Principles
(SOSP '17)* (pp. 153–167). ACM. https://www.microsoft.com/en-us/research/wp-content/uploads/2017/10/Resource-Central-SOSP17.pdf  \n

Microsoft Azure. (2019). *AzurePublicDataset: Microsoft Azure traces* (Version 2) [Data set]. GitHub. https://github.com/Azure/AzurePublicDataset









\n
\n
